Updated results

### 2021-03-22 analysis/rev_prototyping/old/revision_results.Rmd
rev draft

@@

@@

ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts > 20), aes(real_po_percentile_mean)) +
  geom_histogram(bins = 40) +
  facet_wrap(vars(dat, very_small), scales = "free_y", ncol = 2)
```
#### Shape metrics

@@

ggplot(filter(all_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, dat, very_small), scales = "free", ncol = 5) +
  geom_vline(xintercept = c(2.5, 97.5))

@@

jk_cts <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv")) %>%

  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

@@

>>>>

## Datasets

```{r, fig.dim = c(8,6)}

ggplot(all_di, aes(s0, n0, color = Dataset)) +
  geom_point(alpha = .3) +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_viridis_d(end = .8)

@@

ggplot(filter(all_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, dat), scales = "free", ncol = 5) +
  geom_vline(xintercept = c(2.5, 97.5))

@@

ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts > 20), aes(real_po_percentile_mean)) +
  geom_histogram(bins = 40) +
  facet_wrap(vars(dat, very_small), scales = "free_y", ncol = 2)
```

@@

@@

misc_jk_ct <- read.csv(here::here("misc_jk_cts.csv"))
mcdb_jk_ct <- read.csv(here::here("mcbd_jk_cts.csv"))

jk_cts <- read.csv(here::here("analysis", "rev_prototyping", "all_ct_jk.csv")) %>%
  bind_rows(misc_jk_ct) %>%
  bind_rows(mcdb_jk_ct) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

@@

## Gentry

all_di <- all_di %>%
  mutate(low_avgn = n0/s0 < 3)

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes( nsingletons_percentile)) + 
geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y")

all_di %>%
  filter(nparts > 40, dat == "gentry") %>%
  group_by(low_avgn) %>%
  summarize(low_nsingles = mean(nsingletons_percentile < 2.5),
            high_even = mean(simpson_percentile_excl > 97.5),
            high_shannon = mean(shannon_percentile_excl > 97.5),
            low_skew = mean(skew_percentile < 2.5),
            n = dplyr::n())

ggplot(filter(all_di, nparts > 40), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes(nsingletons_median / s0)) +
  geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y", ncol = 1) +
  xlab("Median proportion of rare species")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes(s0, n0, color = low_avgn)) +
  geom_point()

ggplot(filter(all_di, nparts > 40), aes(s0, n0, color = nsingletons_median / s0, shape = Dataset)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  theme(legend.position = "top")

ggplot(filter(all_di, nparts > 40, dat == "gentry"), aes((n0/s0), nsingletons_percentile)) + 
  geom_point() +
  scale_x_log10() +
  geom_hline(yintercept = 2.5)

## Comparison of FIA to other small communities

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

}

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

}

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

fia_di <- filter(sub_di, Dataset == "FIA")
other_di <- filter(sub_di, Dataset != "FIA")

```{r}
filter(sub_di, nparts > 20) %>%
  group_by(Dataset) %>%
  summarize(high_diss = mean(real_po_percentile_excl > 95))

ggplot(sub_di, aes(real_po_percentile_excl)) +
  geom_histogram() +
  facet_wrap(vars(Dataset))
ks.test(fia_di$real_po_percentile_excl, other_di$real_po_percentile_excl)

ggplot(sub_di, aes(sim_pos_from_best)) +
  geom_histogram() +
  facet_wrap(vars(Dataset))

ks.test(fia_di$sim_pos_from_best, other_di$sim_pos_from_best)

```{r, fig.dim = c(7, 4)}

sub_di_shapes_long <- sub_di  %>%
  filter(nparts > 40) %>%
  #mutate(Dataset = ifelse(nparts < 1000, paste0(Dataset, "_small"), Dataset)) %>%
  select(Dataset, skew_percentile, skew_percentile_excl, simpson_percentile, simpson_percentile_excl, shannon_percentile, shannon_percentile_excl, nsingletons_percentile, nsingletons_percentile_excl) %>%
  rename(skew = skew_percentile,
         skewX = skew_percentile_excl,
         simp = simpson_percentile,
         simpX = simpson_percentile_excl,
         shan = shannon_percentile,
         shanX = shannon_percentile_excl,
         nsng = nsingletons_percentile,
         nsngX = nsingletons_percentile_excl) %>%
  mutate(skewM = (skew + skewX) / 2,
         simpM = (simp+ simpX)/2,
         shanM = (shan+ shanX)/2,
         nsngM = (nsng+ nsngX)/2) %>%
  tidyr::pivot_longer(-Dataset, names_to = "metric", values_to = "value") %>%
  mutate(metric_name = substr(metric, 0, 4),
         metric_mod = ifelse(substr(metric, 5, 5) == "X", "excl", ifelse(substr(metric, 5,5) == "M", "mean", "incl")))

ggplot(filter(sub_di_shapes_long, metric_mod == "mean"), aes(value)) +
  geom_histogram(bins = 40) +
  #facet_grid(rows = vars(dat), cols = vars(metric_name), scales = "free", switch = "y") +
  facet_wrap(vars(metric_name, Dataset), scales = "free", ncol = 2) +
  geom_vline(xintercept = c(2.5, 97.5))

sub_di_shapes_high <- sub_di_shapes_long %>%
  filter(metric_mod == "excl") %>%
  group_by(Dataset, metric_name) %>%
  summarize(proportion_high = mean(value > 97.5),
            n = dplyr::n())

sub_di_shapes_high %>%
  tidyr::pivot_wider(id_cols = c(Dataset, n), names_from = metric_name, values_from = proportion_high) %>% mutate(dir = "HIGH")

sub_di_shapes_low <- sub_di_shapes_long %>%
  filter(metric_mod == "incl") %>%
  group_by(Dataset, metric_name) %>%
  summarize(proportion_low = mean(value <2.5))

sub_di_shapes_low %>%
  tidyr::pivot_wider(id_cols = Dataset, names_from = metric_name, values_from = proportion_low) %>%
  mutate(dir = "LOW")

ks.test(fia_di$nsingletons_percentile_excl, other_di$nsingletons_percentile_excl)
ks.test(fia_di$skew_percentile_excl, other_di$skew_percentile_excl)
ks.test(fia_di$simpson_percentile, other_di$simpson_percentile)
ks.test(fia_di$shannon_percentile, other_di$shannon_percentile)
